Open-source landscape for 3D CSEM modelling

### 2020-09-29 manuscript/Manuscript.tex
Tetrahedra plots
Updates from Overleaf

\begin{abstract}
% The (points) are from Mahrer (1993), Geophysics, Some Considerations When Abstracting; DOI: 10.1190/1.58010001.1
% (1) principal objectives and scope of the work
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. However, this has changed over the last few years, and today there exists a selection of different open-source codes which are up to the task. We have a closer look at four different ones, where in all cases at least the user-facing functionality is written in Python.
% (2) methodology
For the numerical comparison we use three models: a simple layered model; a model containing three resistive blocks in a layered background; and a complex, real-world resistivity model. All three models are vertically transverse isotropic. The analysis includes three different mesh-types, simple tensor meshes, octree meshes, and unstructured tetrahedral meshes. Two of the codes use finite volumes and the other two use finite elements.
% (3) results
The comparison shows that all four codes are able to compute the different models to sufficient precision. The biggest difference between codes is actually the discretization and the ease of creating a suitable mesh, particularly as that part often takes much more time than the actual computation. The required runtime and memory varies greatly between the codes. However, these depend mainly on the underlying solver, and not on the actual CSEM codes.
% (4) conclusions
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community at large, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. All our scripts to reproduce the shown results are open-source.
\end{abstract}

@@

The shown examples consider the geophysical problem of marine CSEM. However, the codes could, and many frequently are, applied to very different applications, including any other geophysical exploration in the sea or on land (oil, mining, geothermal, groundwater, environmental) and also non-geophysical tasks. The current main applications of the different codes are mining (\custem), mining and environmental (\simpeg), oil and more recent medical and geothermal (\petgem), and
oil, methane, and deep sea mining (\emg3d).

>>>>

\begin{abstract}
% The (points) are from Mahrer (1993), Geophysics, Some Considerations When Abstracting; DOI: 10.1190/1.58010001.1
% (1) principal objectives and scope of the work
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. However, this has changed over the last few years, and today there exists a selection of different open-source codes which are up to the task. We have a closer look at four different ones, where in all cases at least the user-facing functionality is written in Python.
% (2) methodology
For the numerical comparison we use three models: a simple layered model; a model containing three resistive blocks in a layered background; and a complex, real-world resistivity model. All three models are vertically transverse isotropic. The analysis includes three different mesh-types, simple tensor meshes, octree meshes, and unstructured tetrahedral meshes. Two of the codes use finite volumes and the other two use finite elements.
% (3) results
The comparison shows that all four codes are able to compute the different models to sufficient precision. The biggest difference between codes is actually the discretization and the ease of creating a suitable mesh, particularly as that part often takes much more time than the actual computation. The required runtime and memory varies greatly between the codes. However, these depend mainly on the underlying solver, and not on the actual CSEM codes.
% (4) conclusions
The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community at large, and we invite and encourage the community to make more code, modelling scripts, and results publicly available. All our scripts to reproduce the shown results are open-source.
%
\\~\\ \# \# Alternative \# \# \\
Large-scale modelling of three-dimensional controlled-source electromagnetic (CSEM) responses used to be feasible only for large companies and research consortia. This has changed over the last few years, and today there exists a selection of different open-source codes. Using four different codes in the Python ecosystem we model responses for three increasingly complex models. The verification of the results with semi-analytical solutions for a simple case as well as the validation between the codes and with results from the industry for more complex models show that the open-source codes are able to compute CSEM responses for challenging, large-scale models. Our comparison includes finite-element and finite-volume codes, using structured tensor and octree meshes as well as unstructured tetrahedra meshes. Our results show further that accurate responses can be obtained independently of the chosen method and the chosen mesh. The difficult part, however, is to appropriately discretize the model. The runtime and memory requirements, however, vary greatly between different codes. But we have found that much more time was spent on creating the appropriate input files than running the actual computation. We provide the three models and the corresponding responses of four codes, which can be used for verification of new and existing codes. The collaboration of four code maintainer trying to achieve the same task brought in the end all four codes a significant step further. We hope that these results may be useful for the entire CSEM community at large, and we invite and encourage the community to make more code, modelling scripts, and results publicly available.
\end{abstract}

@@

The shown examples consider the geophysical problem of marine CSEM with resistive bodies. We hope to see in the future similar comparisons for other cases, such as land CSEM with strong topography looking for conductive bodies. The used codes could, and many frequently are, applied to very different applications, including any other geophysical exploration in the sea or on land (oil, mining, geothermal, groundwater, environmental) and also non-geophysical tasks. The current main applications of the different codes are mining (\custem), mining and environmental (\simpeg), oil and more recent medical and geothermal (\petgem), and oil, methane, and deep sea mining (\emg3d).
